---
title: "2nd Assignment EN 2025: Moving vultures"
subtitle: "Proposed Solution"
author: "Tiago A. Marques"
date: "`r format(Sys.Date(), '%d %B %Y')`"
toc: true
number-sections: true
colorlinks: true
format:
  html: 
    code-fold: false
    html-math-method: katex
  pdf: 
    geometry: 
      - top=30mm
      - left=20mm
  docx: default
editor: visual
csl: C:\\Users\\tiago\\Dropbox\\ctr\\mee.csl
bibliography: C:\\Users\\tiago\\Dropbox\\ctr\\MainBibFile.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,cache=FALSE)
options(scipen=999)
```

```{r,echo=FALSE,eval=FALSE}
#if one wants html format
format:
  html:
    toc: true
    html-math-method: katex
    css: styles.css
```

```{r,warning=FALSE,message=FALSE,echo=FALSE,cache=FALSE}
library(readxl)
library(purrr)
library(DT)
```

Use this `.qmd` document (`.qmd` files are quarto documents, for you these work just the same as .rmd files) to see a solution and how your answers differed from my solution.

Select the `.txt` file with your results. You must replace in the assignment of object `s2cf` the name of the `.txt` you handed in instead of `file.txt`. Said file must be in the same folder as this ".qmd" file

Also, substitute below the 11111 assigned to `naluno` with your student number

```{r}
#student to check .txt file
s2cf <- "file.txt"
# student to check number
# number of characters in the file name
ncs <- nchar(s2cf)
naluno <- 11111
```

A few constants are defined.

```{r}
## number of questions in assignment
nq <- 9
```

```{r}
# Weights by question
# 1 more weight than questions
# the first weight is 1, just
# to bring in the student number
# weights must add to 1 (except the first)
# defaults to all weights being equal
weights=c(1,rep(1/nq,nq))
```

Reading required data

```{r}
grifos <- read.csv("training_dataset.csv")
```

The function `getmyres` computes the correct results given a seed, the number of questions, and a dataset (`grifos`, by default). This function is more complicated than what your code might be, this is attempting to be general.

```{r}
getmyres<-function(aluno,nq,mydata = grifos){
# this function takes in as arguments
# aluno: the student number
# nq: the number of questions in the assignment - here nq = 9
# here I compute the correct results
# ----
#first, generate the same data a student would get
set.seed(aluno)
n <- round(runif(1,1000,2000))
# your rows of data
yr <- sample(1:nrow(mydata),size = n,replace=FALSE)
# your data
yd <- mydata[yr,]
# a sample for one of the exercises
n2 <- round(runif(1,10,20))
# your significance values
sig1 <- sample(c(0.1,0.05,0.01,0.001,0.07,0.32),size = 1)
sig2 <- sample(c(0.1,0.05,0.01,0.001,0.07,0.32),size = 1)
# values of the rows for the last exercise
i1 <- round(runif(1,300,500))
i2 <- round(runif(1,501,800))
i3 <- round(runif(1,801,1000))
# ----
#second, get the correct answers given the student number
# all columns with accelerations
cxyz <- 5:304
#columns of x
cx <- 5:104
#columns of y
cy <- 105:204
#columns of z
cz <- 205:304
# number of records
result1 <- nrow(yd)
# for the first record get
# get the x's accelerations
xs1 <- as.numeric(yd[1,cx])
# get the y's accelerations
ys1 <- as.numeric(yd[1,cy])
# get the z's accelerations
zs1 <- as.numeric(yd[1,cz])
# get the correlation between x's and y's accelerations
result2 <- round(cor(xs1,ys1),2)
# get the correlation between x's and z's accelerations
result3 <- round(cor(xs1,zs1),2)
# get the correlation between y's and z's accelerations
result4 <- round(cor(ys1,zs1),2)
# get which one is the maximum
# absolute correlations between pairs of accelerations
absacxyz<- abs(c(result2,result3,result4))
maxis<- which(absacxyz==rep(max(absacxyz),3))
# return the pair with the maximum absolute correlation
result5 <- c("X e Y","X e Z","Y e Z")[maxis]
# check if it is a positive or a negative correlation
result6 <- ifelse(c(result2,result3,result4)[maxis]>0,"positive","negative")
# get the value of the t.test
# comparing the xs and the ys for the first n2 rows
# and note it is a paired test
# the unlist function is not strictly necessary but might avoid issues
mytest <- t.test(unlist(yd[1:n2,cx]),unlist(yd[1:n2,cy]),paired = TRUE)
# get the value of the test statistic 
result7 <- round(as.numeric(mytest$statistic),2)
# make your decision at the "sig1" level of significance
decision <- ifelse(mytest$p.value<sig1,"reject H0","do not reject H0")
# store decision
result8 <- decision
# format data to compare the absolute accelerations across 3 rows of data
# those are rows i1, i2 and i3
myrows <- c(paste0("L",i1),paste0("L",i2),paste0("L",i3))
# there will be 300 records for each row
# and we want the absolute accelerations
# so need to use the fucntion "abs"
dataL1 <- as.numeric(yd[i1,cxyz])
dataL2 <- as.numeric(yd[i2,cxyz])
dataL3 <- as.numeric(yd[i3,cxyz])
accels.i1i2i3 <- data.frame(
  row=as.factor(rep(myrows,each=300)),
  accel= c(dataL1,dataL2,dataL3))
# #note the code above is complicated but one could do this in a simpler way.
# di1 <- as.numeric(yd[i1,cxyz])
# di2 <- as.numeric(yd[i2,cxyz])
# di3 <- as.numeric(yd[i3,cxyz])
# grupo <- rep(c("L1","L2","L3"),each=300)
# accels.1 <- data.frame(row=
#     grupo,
#     accel= c(di1,di2,di3))
# do the Kruskall Wallis test, take the corresponding p-value
kwtestp <- kruskal.test(formula=accel~row,data=accels.i1i2i3)$p.value
# record your decision
decision <- ifelse(kwtestp<sig2,"reject H0","do not reject H0")
result9 <- decision
#store answers in the correct format
  myanswers <- list(aluno)
  for(i in 1:nq){
    myanswers[[i+1]]<-
      eval(parse(text=paste0("result",i)))
  }
  return(myanswers)
}
```

Your results before manual tweaking

```{r}
# get student results
studentres <- dget(file=s2cf)
studentres
```

and after manual tweaking (in case you did not use a numeric but you got the right test I consider the answer to be correct)

```{r}
# if they put all the blurb from the correct test I consider it incorrect
# the guidelines are clear
# this must be a numerical value
if(class(studentres[[8]])=="htest") studentres[[8]]<-0
# if they gave me the correct test statistic,
# even if not numeric, I consider it correct
studentres[[8]] <- as.numeric(studentres [[8]])
studentres
```

Correct results

```{r}
# get true results for student i
# this is the hard stuff, needs to be tweaked each time
# by creating a bespoke code inside function "getmyres"
trueres <- getmyres(aluno=studentres[[1]],nq=nq)
trueres
```

Check which results are correct

```{r}
correctres<- purrr::map2_lgl(studentres, trueres, ~ isTRUE(all.equal(.x, .y, tolerance = 1e-7)))[-1]
correctres
```

Your grade is `r round(sum(correctres)/length(correctres),3)`.

The incorrect results correspond to question(s)

```{r}
which(!correctres)
```

the results that are different are shown below, your results

```{r}
studentres[c(FALSE,!correctres)]
```

then what they should be

```{r}
trueres[c(FALSE,!correctres)]
```

For easier comparison, the top row is your result, and the bottom row the correct result, for the questions you got incorrectly

```{r}
rbind(studentres[c(FALSE,!correctres)],trueres[c(FALSE,!correctres)])
```
